\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage{sigsam, amsmath}
\newtheorem{theorem}{Theorem}

\usepackage[ruled,vlined,linesnumbered]{algorithm2e}
\SetEndCharOfAlgoLine{}

% additional commands
\newcommand{\F}{\mathbb{F}}
\newcommand{\FF}{\F}
\renewcommand{\O}{\mathcal{O}}
\newcommand{\Ell}{\mathrm{Ell}}
\newcommand{\Cl}{\mathcal{C}}
\newcommand{\ms}{\mathrm{\ ms}}
\newcommand{\s}{\mathrm{\ s}}
\newcommand{\frakell}{\mathfrak{l}}
\newcommand{\Walk}{\textsc{Walk}}
\newcommand{\Step}{\textsc{Step}}

% leave as is
\issue{TBA}
\articlehead{TBA}
\titlehead{Isogeny-based cryptography in Julia/Nemo: a case study}
\authorhead{Jean Kieffer, Luca De Feo}
\setcounter{page}{1}

\begin{document}

\title{Isogeny-based cryptography in Julia/Nemo: a case study}

\author{Jean Kieffer \\
\'Ecole normale sup\'erieure de Paris \& INRIA, Universit\'e Paris-Saclay \\
\url{jean.kieffer@ens.fr}
\and
Luca De Feo \\
Universit\'e de Versailles \& INRIA, Universit\'e Paris-Saclay \\
\url{luca.defeo@polytechnique.edu}}

\maketitle

\begin{abstract}
The Couveignes--Rostovtsev--Stolbunov key-exchange protocol based on
    isogenies of elliptic curves is of interest because it may resist
    quantum attacks, but its efficient implementation remains a challenge.
    We briefly present the computations involved, and efficient
    algorithms to achieve the critical steps, with timing results for our implementations in Sage and Julia/Nemo.
\end{abstract}

\section{Introduction}
\label{sec:intro}

In \cite{RS}, Rostovtsev and Stolbunov proposed using isogenies between
ordinary elliptic curves over finite fields to build key-exchange and
public-key encryption schemes, rediscovering Couveignes' idea
\cite{Couveignes}. These schemes are based on the simply transitive action
of the abelian group $\Cl(\O)$, the class group of an order \(\O\) in a
quadratic field, on $\Ell_t(\O)$, the set of elliptic curves over a
(prime) finite field $\F_p$ with Frobenius trace $t$ and complex
multiplication by the order $\O$. Elements in $\Cl(\O)$ are represented
as products of ideals of prime norm, up to a certain bound. 
Each of these ideals \(\frakell\) is represented by its norm \(\ell\) and a
\emph{Frobenius eigenvalue} \(v_\ell\),
which is one of the two roots of \(X^2 - tX + p \pmod{\ell}\).
Computing the action of \(\Cl(\O)\) then amounts to following a path in
the so-called \emph{isogeny graph}, whose vertices are curves in $\Ell_t(\O)$ and edges are isogenies corresponding to ideals. 

The cryptosystem parameters are
a set \(L\) of prime ideals in \(\O\),
an exponent bound \(K\),
and a ``base point'' \(E_0\) in \(\Ell_t(\O)\).
Key exchange works as follows:
Alice samples a secret \(k_A\) from \([0,K]^L\)
and sends \(E_A = \Walk(E_0,k_A)\) to Bob;
Bob samples a secret \(k_B\) from \([0,K]^L\)
and sends \(E_B = \Walk(E_0,k_B)\) to Alice.
Their shared secret is \(E_{AB} \cong \Walk(E_A,k_B) \cong \Walk(E_B,k_A)\).
Each \(\Walk\) (Algorithm~\ref{alg:walk})
is a series of \(\Step\)s
for each of the various ideals \(\frakell\) in \(L\);
we describe two strategies for implementing \(\Step\) below,
comparing implementations in Sage \cite{Sage} and Nemo \cite{Nemo}. 

The choice of \(L\) is critical,
because the security of the scheme grows exponentially with \(\#L\)
(and only polynomially in \(K\)),
but the efficiency of each \(\Step\)
depends on the particular \(\frakell\) in~\(L\).
For our experiments,
we chose parameters aiming at the 128-bit security level
(comparable with contemporary cryptosystems such as AES)
against the best known classical attacks:
%To fix ideas, let us give an example of parameters achieving security comparable with contemporary cryptosystems, against the best known classical attacks. 
we take
$p~=~2^{500} + 55$, 
$K = 2$, 
$E_0: y^2 = x^3 + 5x^2 + x$,
and $L$ a set of 115 convenient primes with norms less than $1400$.
%(out of 222),
For each \(\frakell\) in \(L\),
the classical modular polynomial \(\Phi_\ell(X,Y)\)
is precomputed.

\begin{algorithm}
    \caption{{\sc Walk}: a walk in the isogeny graph}
    \label{alg:walk}
    \KwIn{An elliptic curve \(E\) in \(\mathrm{Ell}_t(\O)\)
    and a vector \(k = (k_\frakell)_{\frakell \in L} \in [0,K]^L\)
    }
    \KwOut{An elliptic curve \(E'\) such that \(E \to E'\)
    is an isogeny corresponding to \(\prod_{\frakell\in
    L}\frakell^{k_\frakell}\)}
    %
    \For{\(\frakell \in L\)}{
        \lFor{\(0 \le i < k_\frakell\)}{%
            \label{alg:walk:step-call}
            \(E \gets \textsc{Step}(E,\frakell)\)
        }
    }
    \Return{\(E\)}
\end{algorithm}


%Note that we cannot use all primes, but only about half of them: these primes are usually called \emph{Elkies primes} in point counting. One could try to find special curves, for which more primes could are Elkies. It is a major issue since the security of the scheme grows exponentially with the length of $L$ and only polynomially with $K$, and we return to this question later on.

\section{A first solution}

\begin{algorithm}
    \caption{{\sc Step}: a single step in the isogeny graph, using
    Bostan--Morain--Salvy--Schost}
    \label{alg:step}
    \KwIn{%
        $E\in \Ell_t(\O)$, 
        a prime ideal \(\frakell \leftrightarrow (\ell,v_\ell)\)
    }
    \KwOut{%
        $E'$ such that \(E \to E'\) is the edge in the isogeny graph
        corresponding to \(\frakell\)
    }
    %
    Evaluate \(F(X) = \Phi_\ell(X,j(E))\)
    \tcp*{\(j\) is the \(j\)-invariant; \(\Phi_\ell\) is precomputed}
    %
    \(\{j_1,j_2\} \gets \textsc{Roots}(F,\F_p)\);
    compute curves $E_1$, $E_2$ such that \(j_i = j(E_i)\)
    \label{alg:step:roots}
    \;
    %
    \(S_1 \gets \textsc{KernelOfIsogeny}(E,E_1,\ell)\)
    \label{alg:step:kernel}
    \;
    %
    \lIf{\((x^p \pmod{S_1}) = ([v_\ell]_*x \pmod{S_1})\)}{%
        \Return{\(E_1\)}
    }
    \lElse{%
        \Return{\(E_2\)}
    }
\end{algorithm}

Algorithm~\ref{alg:step}
is our first and most general implementation of \(\Step\).
In Line~\ref{alg:step:kernel},
we use the Bostan--Morain--Salvy--Schost algorithm \cite{BMSS}. 
This involves finding a rational function satisfying a differential
equation, which is solved using a Newton iteration,
before recovering 
the polynomials we want using the Berlekamp--Massey algorithm. 
The cost is dominated by finding the roots of $\Phi_\ell(X, j(E))$, 
which has degree $\ell + 1$, 
in Line~\ref{alg:step:roots}. 
For this, we use the Cantor--Zassenhaus algorithm, performing equal-degree factorisation only in degree 1.
The most costly step is computing $X^p\pmod{F}$.
We used Sage
(which relies on NTL \cite{NTL}, PARI \cite{Pari} and GMP \cite{GMP}
here)
and Nemo built-ins for this common computation, 
giving the following results
(all timings are from a 3.20GHz Intel Xeon workstation).
\[
    \begin{array}{c|c|c|c|c|c|c|c|c|c}
        \deg F & 2 & 5 & 10 & 20 & 50 & 100 & 200 & 500 & 1000\\
        \hline
        \text{Sage} & 40 \ms & 40 \ms & 55\ms & 70 \ms & 155\ms & 510\ms & 1.0 \s & 2.2 \s & 5.3 \s\\
        \hline
        \text{Nemo} & 3.5 \ms &  14 \ms & 38 \ms & 90 \ms & 220 \ms & 460 \ms & 980 \ms & 2.5 \s & 5.5 \s
    \end{array}
\]

Sage also spends additional time in precomputations (as much as 7 minutes in the last example). While this may prove efficient when a lot of calculations are to be done in the same ring, this is not the case in our setting. Using the parameters fixed in~\S\ref{sec:intro},
the total computing time is approximately 7 minutes for each
\(\Walk\), with about 20 seconds spent on each of the largest primes.


\section{Faster $\ell$-isogenies with torsion points}

\begin{algorithm}
    \caption{\(\Step\) using torsion points}
    \label{alg:torsion}
    \KwIn{$E\in \Ell_t(\O)$ and a prime ideal \(\frakell \leftrightarrow
        (\ell,v_\ell)\) such that \(\ell\mid \#E(\FF_{p^d})\)
        for a small \(d\)
    }
    \KwOut{%
        \(E'\) such that \(E \to E'\) is the edge in the isogeny graph
        corresponding to \(\frakell\)
    }
    %
    \Repeat{\(Q \not= 0\) \tcp*[f]{\(Q\) is a point of exact order \(\ell\)}}{
        \(P \gets \textsc{Random}(E(\FF_{p^d}))\)
        \;
        %
        \(Q \gets [C_d/\ell]P\)
        where \(C_d = \#E(\FF_{p^d})\)
        \tcp*{\(C_d\) is a global constant}
        \label{alg:torsion:scalarmult}
    }
    %
    \(S \gets \{[i]Q: 1 \le i \le (\ell-1)/2\}\)
    \tcp*{Use \((\ell-1)/2-1\) repeated additions}
    %
    \Return{\(\textsc{QuotientCurve}(E,S)\)}
    \tcp*{Use Vélu's formulae}
\end{algorithm}


If $E$ has points of order $\ell$
defined over \(\FF_p\), or a small-degree extension of~\(\FF_p\),
then we may replace Algorithm~\ref{alg:step}
with a faster method, Algorithm~\ref{alg:torsion}.
The costly part here is the scalar multiplication in
Line~\ref{alg:torsion:scalarmult}, assuming $\ell$ is not too large;
typically \(C_d/\ell\sim p^d/\ell\).
In order to optimize the number of $\F_p$-operations, we implemented it
using Montgomery curves and \(x\)-line arithmetic~\cite{Montgomery}, 
achieving dramatic performance gains compared to
Algorithm~\ref{alg:step}.
As elliptic curves are not yet supported in Nemo, we implemented our own Montgomery arithmetic. In Sage we use the built-in scalar multiplication. 
Our timings (again on a 3.20GHz Intel Xeon workstation) are as follows:
\[
    \begin{array}{c|c|c|c|c|c|c|c|c|c}
        d & 1 & 3 & 4 & 5 & 7 & 8 & 9 & 11 & 50\\
        \hline
        \text{Sage} & 38 \ms & 190 \ms & 270 \ms & 320 \ms & 470 \ms & 530 \ms & 610 \ms & 780 \ms & 5.7 \s\\
        \hline
        \text{Nemo} & 7 \ms & 23 \ms & 30 \ms & 38 \ms & 55 \ms & 65 \ms & 75 \ms & 130 \ms & 620\ms
    \end{array}
\]
Further experiments show that Julia's compilation and fast arithmetic on Montgomery curves are equally reponsible for these gains.

To fully exploit this speed-up, we have to find an \(E_0/\F_p\)
such that many small primes divide \(\#E_0(\FF_{p^d})\) for small~\(d\). 
What can be achieved in a reasonable amount of time, and what speed-up we can obtain, seem interesting questions. 


\begin{thebibliography}{99}

\bibitem{BMSS}
A. Bostan, F. Morain, B. Salvy, \'E. Schost.
\newblock Fast algorithms for computing isogenies between elliptic curves.
\newblock {\em Math.~Comput.} 77:263:1755--1778 (2008).

\bibitem{Couveignes}
J.-M. Couveignes.
\newblock Hard homogeneous spaces.
\newblock Preprint (2006).

\bibitem{Nemo}
C. Fieker, W. Hart, T. Hofmann, F. Johansson et. al.
\newblock the Nemo computer algebra package, version 0.6.0, \url{http://nemocas.org}.

\bibitem{GMP}
T. Granlund and the GMP development team.
\newblock The GNU Multi Precision Arithmetic Library, \url{http://gmplib.org/}

\bibitem{Montgomery}
P. L. Montgomery.
\newblock Speeding the Pollard and Elliptic Curve Methods of Factorization.
\newblock {\em Math.~Comput.} 47:177:243--264 (1987).

\bibitem{Pari}
The PARI~Group.
\newblock PARI/GP version {\tt 2.9.0}, Univ. Bordeaux, 2016, \url{http://pari.math.u-bordeaux.fr/}.

\bibitem{RS}
A. Rostovtsev, A. Stolbunov.
\newblock Public-key cryptosystem based on isogenies.

\bibitem{Sage}
The Sage Developers.
\newblock {S}ageMath, the {S}age {M}athematics {S}oftware {S}ystem ({V}ersion
  7.5.1),
\newblock 2017, \url{http://www.sagemath.org}

\bibitem{NTL}
V. Shoup.
\newblock NTL : a library for doing number theory.
\newblock \url{http://www.shoup.net/ntl/}

%\bibitem{Waterhouse}
%W. C. Waterhouse.
%\newblock Abelian varieties over finite fields.
%\newblock {\em Ann. sci. \'E.N.S.} 2:4:521--560, 1969.

%\bibitem{Attack}
%S. Galbraith.
%\newblock Constructing isogenies between elliptic curves over finite fields.
%\newblock {\em LMS J. Comput. Math.} 2:118--138, 1999.

%\bibitem{Kohel}
%D. Kohel.
%\newblock Endomorphism rings of elliptic curves over finite fields.

%\bibitem{Flint}
%W. Hart, F. Johansson, S. Pancratz.
%\newblock {FLINT}: {F}ast {L}ibrary for {N}umber {T}heory, 2013,
%\newblock Version 2.4.0, \url{http://flintlib.org}

\end{thebibliography}

\end{document}
